### 2020-05-07 doc/manuscript/ms.Rmd
intro drafted

---
title: "Testing the thermal buffer of mire groundwater"
author: "Eduardo Fernández-Pascual, University of Oviedo (fernandezpeduardo@uniovi.es)"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
output:
  bookdown::html_document2
bibliography: refs.bib
biblio-style: apalike
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(here); library(kableExtra)
load(here("results", "logs.RData"))
source(here("src", "analysis.R"))
```

# Introduction

Climate change is already affecting global biodiversity patterns (IPCC 2014). Among other trends, there has been an expansion of drylands (Huang, J., Yu, H., Guan, X. et al. Accelerated dryland expansion under climate change), an increase of disturbations in forests (Seidl, R., Thom, D., Kautz, M. et al. Forest disturbances under climate change. Nature Clim Change 7, 395–402 (2017). https://doi.org/10.1038/nclimate3303) and ocean acidification (Coral Reef Ecosystems under Climate Change and Ocean Acidification). As (M. Horsák et al. / Science of the Total Environment 634 (2018) 355–365) point out, some characteristics of ecosystems that make them especially vulnerable to climate change are (i) a preponderance of species that evolved under a cold climate, (ii) low productivity caused by nutrient limitation, which makes them sensitive to increased nutrient
cycling caused by warming (Cornelissen et al., 2007), and (iii) a scatterred distribution that limits species dispersal and migration (Pearson and Dawson, 2005). These characteristics are found in temperate peatlands such as mires, permanent semi-terrestrial wetlands where soils remain waterlogged but not inundated during most of the year (Wheeler and Proctor, 2000). Mires are priority habitats for conservation that harbour a high number of endangered species (Bergamini et al. 2009) and support highly-adapted flora in spatially reduced areas (Grootjans et al. 2006, Jiménez-Alfaro et al. 2013). Habitat distribution models have predicted a loss of mire surface in face of climate change (Essl et al. 2012).

Recent research has focused on the importance of groundwater as a regulator of mire microclimate. As azonal habitats whose occurence is driven by local edaphic conditions, locally obtained
thermal data are hence crucial for any prediction about the impact of climatically-induced warming (Breckle 2002). In Europe, fens are distributed from the Iberian Peninsula to Fennoscandia and from low valleys to the alpine belt, but nonetheless the constituent and dominant species remain relatively homogeneous (Jiménez-Alfaro et al. 2014). An apparent climatic resilience of fen species is further suggested by the persistence of relicts in their distribution limits (Amon et al. 2002, Topić and Stančić 2006, Hájek et al. 2009, Jiménez-Alfaro et al. 2012). A possible cause for this climatic resilience may be the buffer effect of groundwater on soil temperature, reducing the expected differences of the realized niche in diverse climatic regions (Ellenberg 1988). Indeed, the cool waters from underground aquifers that saturate fens are expected to keep temperatures warmer during cold periods and vice versa (Ellenberg 1988, Geiger et al. 2009). As root-zone temperature is known to be a major determinant of plant ecophysiology (Körner and Paulsen 2004), this buffer effect might allow fen species to regenerate and live along a wider range of air temperatures than they could otherwise. Soil temperature measurements have become available for mires of North America (Raney 2014), Western Europe (Fernández-Pascual et al. 2015), Central Europe (Horsak 2018). The comparison of said temperatures with air temperatures derived from models has shown that mire soil temperature is warmer in winter and cooler in summer, thus indicating the existence of a thermal buffer by groundwater (FerPas, Horsak). This effect has been linked to the composition of flora and fauna (Horsak 2018, Shecnkova 2020), growth of mire trees /Raney 2016) and the role of mires as glacial refugia (JimAlf 2016, Dite 2017). It is evident that the thermal buffer of groundwater will play a profound role in the response of these habitats to climate change, and needs to be studied in more depth.

To my knowledge, the soil thermal buffer of mires has not yet been compared to adjacent dry soild with temperatures measured in situ. This article provides the first record of the thermal buffer against surrounding non-mire areas, based on soil temperatures recorded during a period of five years. Specifically, this article test the hypothesis that temperatures in the mire soil will be warmer in winter, colder in summer and less variable than temperatures in adjacent dry soils.

# Materials and methods

This study took place in the temperate oceanic region of north-western Spain (43º N, 5º W). Local geography and climate are dominated by the Cantabrian Mountains (> 1500 m above sea level), which run parallel to the coast and trap the prevailing NW Atlantic winds. The resulting humid climate harbours the south-western limit of mire communities in Europe [@RN2982; @RN3246]. Rain-fed raised bogs and acid valley mires can be found from the coast to just below the treeline, in poorly drained valleys and former glacial lakes. Glacial lakes undergoing silting develop transition mires and quaking bogs communities in the water-to-land transition. Spring fens appear in the mountains above 1000 m; they range from soft-water poor fens on acid bedrocks, to alkaline calcareous fens on limestone.

A selection of `r length(unique(temperatures$Logger))/2` mire sites was made, representing the regional elevation gradient of mire vegetation and the different mire types (Table \@ref(tab:table1))). In each site, two dataloggers (M-Log5W, GeoPrecision, Ettlingen, Germany) were buried at a depth of 5 cm below the upper layer of the soil. They were programmed to record temperature on an hourly basis. One datalogger was placed in a flat waterlogged spot, within the mire. The other dataloggers was placed in the close vicinity, but in a flat and dry area outside the mire. The vegetation was always either mire or pasture, with no shrubs, trees or any other landscape features shading the measuring points. The dataloggers were left on site for approximately five years, after which they were retrieved and their records downloaded. At the moment of retrieval, the internal clock of all the dataloggers had not deviated for more than four hours.

Some steps were taken to clean the logs. First, the records from the first week after installation were removed, to account for the own installation process and the settling of the soils. Second, and because some of the dataloggers had failed at different points in time, a selection was made for each site to keep only time series with records both for the dry and waterlogged points. Afterwards four bioclimatic variables were calculated for each datalogger:

* The mean diurnal range; i.e. the average for the whole period of the daily differences between the maximum and the minimum temperatures recorded in the day.
* The maximum temperature of the warmest month; i.e. the average of the daily maximum records, for the warmst month.
* The minimum temperature of the codlest month; i.e. the average of the daily minimum records, for the coldest month.
* The annual range; i.e. the difference between the maximum temperature of the warmest month and the minimum temperature of the coldest month.

To test if the differences between the dry and the waterlogged points of each site was significant, paired t-tests were used. Tests were one-tailed, according to the following hypotheses: the dry site would have a higher diurnal range, a higher maximum temperature, a lower minimum temperature, and a higher annual range.

# Results

The dataloggers recorded temperatures for five years in five of the sites, four years in two, and two years in one (Fig. \@ref(fig:fig1)). A visual inspection of the time series shows more homogeneous temperatures in the waterlogged time series, which are almost always encompassed within the dry time series. The bioclimatic variables (Table \@ref(tab:table2)) support this impression. The maximum temperatures of the warmest months are usually higher in the dry measuring points, by more than 6 ºC in some sites, both at low (El Molinucu, La Malva) and high (La Recoleta) elevations. The opposite is true for the minimums of the coldest months, in which case the temperature is generally colder in the dry points, although the difference is less pronounced than for the maximums. The mean diurnal range of temperatures is wider in the dry sites, as is annual temperature range; and again the effect is larger in the low sites and in La Recoleta. The t-tests supported the original hypotheses; namely waterlogged measuring points had (a) smaller diurnal fluctuactions (t = `r ttests %>% filter(Trait == "Mean.Diurnal.Range") %>% pull(t.statistic) %>% round(2)`, p = `r ttests %>% filter(Trait == "Mean.Diurnal.Range") %>% pull(t.p.value) %>% round(3)`, effect size = `r ttests %>% filter(Trait == "Mean.Diurnal.Range") %>% pull(t.estimate) %>% round(2)` ºC); (b) lower maximums (t = `r ttests %>% filter(Trait == "Max.Temperature.of.Warmest.Month") %>% pull(t.statistic) %>% round(2)`, p = `r ttests %>% filter(Trait == "Max.Temperature.of.Warmest.Month") %>% pull(t.p.value) %>% round(3)`, effect size = `r ttests %>% filter(Trait == "Max.Temperature.of.Warmest.Month") %>% pull(t.estimate) %>% round(2)` ºC); (c) higher minimums (t = `r ttests %>% filter(Trait == "Min.Temperature.of.Coldest.Month") %>% pull(t.statistic) %>% round(2)`, p = `r ttests %>% filter(Trait == "Min.Temperature.of.Coldest.Month") %>% pull(t.p.value) %>% round(3)`, effect size = `r ttests %>% filter(Trait == "Min.Temperature.of.Coldest.Month") %>% pull(t.estimate) %>% round(2)` ºC), and (d) smaller annual fluctuations (t = `r ttests %>% filter(Trait == "Temperature.Annual.Range") %>% pull(t.statistic) %>% round(2)`, p = `r ttests %>% filter(Trait == "Temperature.Annual.Range") %>% pull(t.p.value) %>% round(3)`, effect size = `r ttests %>% filter(Trait == "Temperature.Annual.Range") %>% pull(t.estimate) %>% round(2)` ºC).

# Discussion

# Tables

```{r table1, echo = FALSE}
table1 %>% kable(caption = "Mire sites included in this study, indicating the type of fen, the elevation, coordinates, and length of the temperature recording period.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

```{r table2, echo = FALSE}
table2 %>% kable(caption = "Bioclimatic variables calculated for the mire sites, using the hourly temperature records.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

# Figures

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Hourly soil temperature records at the mire sites. The blue series was recorded within the mire, in a waterloged area. The red series was recorded in a neighbouring dry area. Dataloggers were buried at 5 cm depth.", fig.align = "center", fig.width=3, fig.height=8}
plot1
```

# References