### 2020-08-29 doc/response-reviewers.Rmd
Revert "1st revision 🍋"
This reverts commit cf4453643d0bd19f444758d41499e7b905a136d8.

---
title: "Ref. JOC-20-0599 - Response to reviewers"
author: "Eduardo Fernández Pascual & Eva Correia-Álvarez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Response to the Editor

# Response to Referee 1

# Response to Referee 2

Dear Dr Fernández-Pascual

We recognise that the impact of the COVID-19 pandemic may affect your ability to return your revised manuscript to us within the requested timeframe. If this is the case, please let us know.

Manuscript # JOC-20-0599 entitled "Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils" which you submitted to the International Journal of Climatology, has been reviewed.  The comments of the referee(s), all of whom are leading international experts in this field, are included with this letter.

The reviewers believe that the paper has potential to be published in the International Journal of Climatology, but major revisions are required. Revisions in this category may involve substantial text changes, recalculations or new analyses in addition to more minor clarifications and corrections.  In your revisions, please address all of the points made by each reviewer, amend the manuscript accordingly, and include a point-by-point response to each reviewer comment with your response.

Both reviewers found this to be an interesting study. There is room to expand on your results/discussion as highlighted by both reviewers to improve upon the rigor and impact of this work.

One upshot of the study is that mire regions might be climate refugia. Figure 3 gets at this point for summer Tmax spatially. I wonder if one can also show how soil temperatures may vary at paired sites during exceptionally warm daily temperatures during the warm season (e.g., complementary temporal analysis).

Because the manuscript needs major revisions, I will send the revised manuscript and your responses back to some or all of the reviewers for further review. Please note that submitting a revision of your manuscript does not guarantee eventual acceptance. I would like to receive the revised manuscript as soon as is convenient, but by 24-Nov-2020 at the latest.

This should be sufficient time given the nature of the revisions, but please communicate with me via email if you need an extension of the deadline.

If the reviewers have suggested that the language of the current version of your manuscript does not allow a thorough evaluation, we recommend that you have your paper professionally edited for English language by a service such as Wiley’s at http://wileyeditingservices.com. Please note that while this service will greatly improve the readability of your paper, it does not guarantee acceptance of your paper by the journal.

Once again, thank you for submitting your manuscript to the International Journal of Climatology and I look forward to  receiving your revision.

Sincerely,

Dr John Abatzoglou
Associate Editor, International Journal of Climatology
jabatzoglou@ucmerced.edu

***************************************************************************
Instructions for submitting your revised manuscript:

To submit your revised manuscript, log on to http://mc.manuscriptcentral.com/joc and click on ‘Author Centre’. Click the link in the ‘Manuscripts with Decisions’ list. Click on the ‘create a revision’ link for the appropriate manuscript.

Please note:
-Once you begin to submit the revision, the 'create a revision' link disappears and the manuscript moves into your Revised Manuscripts in Drafts list. The option reappears if you delete the revision before it is actually submitted.

Should a revised manuscript be accepted, the files will be sent electronically to the Production Department. It is thus very important that the files uploaded at revision are in their final format and of sufficient quality for publication. At revision, you should provide:
a) The manuscript text in Word format (all tracked changes removed). This file should be text only, should not contain figures and must not exceed 1MB in size. Tables must not be embedded, but provided in editable format.
b) Figures as high-resolution TIFF, EPS or PDF files; colour in CMYK format.
c) Supporting Information files as required. Please note that the maximum file size for each Supporting Information file is 20 MB.
d) A copy of the text file with changes highlighted.
e) A letter of response to the reviewers' comments, detailing point-by-point the changes made and giving reasons for those not made.

Revised manuscripts must be in their final form when submitted. The proofs received later are for correction of typographical errors only. They should not be used for final changes to articles; such changes must be made to the manuscript before it goes to Production

***************************************************************************

Due Date of Revised Submission: 24-Nov-2020

***************************************************************************

Referee(s)' Comments to Author:

Reviewer: 1

Comments to the Author
This is a short and simple study comparing 5cm soil temperatures in mire habitats with adjacent drier microhabitats. It could be of interest to some readers, although it could use some polish and set in a broader context.

Specific comments:

L15 – I think a yes/no type hypothesis about climate variability is not very interesting given we already have a priori expectations in this regard (e.g. L61-65, Ashcroft & Gollan 2013 Agric. For. Meteorol. 176:77-89. I would focus more on a quantitative question like: how different are they? What are you finding that is different from what we already know?

L25 – What do you mean by failed to predict? E.g. r2 < 0.5? RMSE > 2oC? Spatial pattern wrong? Consistent bias?

L42 – “has been”

L88-94: Like L15 above, we already have strong evidence that soil moisture acts as a buffer, so testing yes/no hypothesis about this make the research more trivial that it could be. I would focus more on how different they are etc.

L142 – significantly? P<0.05?

L155 – This is more like a results section with actual figures. Would be good to see 95% CI for the effect sizes. i.e. instead of just effect size = -2.29oC, you would give a range like -1.5 to -3.0.

Fig. 3 – Make the scales on the y and x axes the same and add a 1:1 line. R2 tells you the correlation between soil and air temps but not the bias. You could have a high r2 and still have a high RMSE indicating poor predictions.

L160-165 – building on above comment, note both the r2 and RMSE to capture both correlation and bias. These results will be affected by a number of factors: 1) the soil moisture effect you are interested in 2) the height of observation (2m air temperatures v 5cm soil temperatures) 3) the temporal scale of obvservation (the 1-5 years of your study is different to Chelsa) 4) The spatial scale – does Chelsa reflect the actual elevation of your sites.

L167 – “Quantify the size of the thermal buffer effect” rather than “prove the existence”. We already know it exists.

Discussion: Where results are similar or different from other studies it would be nice to delve deeper into why. Were the observations at the same depth? E.g. We expect diurnal range to decrease with depth so studies at different depths are not necessarily comparable. What do the results tell us more generally about mires elsewhere, or where sites may vary along a moisture gradient rather than just be mire or non-mire?

198-208 – Don’t read too much into individual sites. When you only have 8 sites each will be unique in one way or another and we expect some results to be different. What we are interested in is trends that are consistent across all 8 sites.

Reviewer: 2

Comments to the Author
Very interesting datasets. More rigor in describing how analyses were performed and digging a little deeper into the results is needed. Peatland terms vary somewhat from N. America to Europe / Eurasia. Clearly defining hydrological terms like "mire" will be useful - is this a broad term for all peatlands or simply groundwater fed ones?  You describe a groundwater buffer effect, but include wetlands in your study whose hydrology may be dominated moreso by precipitation inputs (bogs, raised bogs) . While it may not be practical to fully describe source hydrology to your sites from in situ measurements, some more information on hydrology of these wetlands would be useful. Does a gradient in groundwater connection strength exist among your sites?

A general comment is, you would expect to see a thermal buffer effect just from the presence of water at all compared to dry sites. Your data indicate that none of the waterlogged sample locations dried out, thus removing the thermal buffer effect during mreasurement (Figure 1). Do you have a hypothesis for what would happen if they dried based on your data? For your low elevation sites, it suggests any locations drying might be really susceptible to warming.

Is it possible to compare/contrast sites based on sites with weaker vs. stronger groundwater inputs in your study sites? Yes there is an elevation gradient present, but do differences in source hydrology, stronger groundwater inputs explain the really pronounced buffering of some sites vs. others irrespective of elevation? Your methods section is really short and does not provide extensive detail on how you performed your analysis, what statistical platform you used, elsewhere R is referenced, but not in the methods directly.

Your study might benefit from using a mixed model instead of a simple linear model. I could be convinced a more sophisticated analytical structure like linear mixed modeling (site random effect) is not needed - especially if you expect differences related to hydrology among sites, but those differences if they exist you may want to provide some justification for. If you think your sites have similar hydro-logical characteristics, then a site random effect grouping wet and dry locations by site may make sense to control for noise in the dataset from microclimate factors such as aspect etc. Please at include some justification as to why a mixed modeling approach is not needed or replace the simple linear model with a mixed model.

It looks as though some of the temperature variability in your dataset is due to differences in hydrology among sites. If you suspect that is the case as well, then you can pretty easily justify the existing linear model analytical structure, but need more interpretation of hydrological differences in your results and discussion.

Do any patterns emerge if you code sites by habitat type?  Please include the habitat type labels presented in Table 1 in the figure panels for Figure 1.

Can you clarify if dry sites are in fact adjacent dry uplands or dry raised hummocks (high spots) within the wetlands? A figure with a site-level inset map would also convey this information and would benefit the paper.

It would be worth describing the >15 degree C differences between wet vs dry sites at low elevation (La Malva). That is very biologically significant! You should put that into perspective for the readers, that level of difference greatly exceeds IPCC projections for climate change in this region... Huge differences within site. Please provide some descriptive statistics on peak differences observed at some of your sites.

This is a neat study, there seems to be some more interesting angles you can explore/present in your discussion that will make your paper more interesting to a wider array of readers.

# PDF

The authors have collected compelling data in an understudied field. The study is relevant to biodiversity under climate change, and to assumptions of global carbon dynamics. The data collected has a lot of merit and pretty easily stands on its own without much help. However, additional descriptions of methodologies are needed. There are some interesting results here, and a nice temporal record from which to draw conclusions. The authors focus on the waterlogged nature of the soils – but the source of hydrology and how closely coupled groundwater in mires is linked with atmospheric conditions may have some bearing on whether these systems remain buffered from effects of atmospheric warming. More robust discussion on the atmospheric coupling of hydrology in these systems tied to other literature is needed. Do groundwater-fed mires persist in drying climates? The paper may benefit from discussion on changes in climate in the study region. Is this study region getting drier? What effect is that expected to have on the thermal buffer effect described? Some descriptions of the latitudinal and elevational distribution of mires in Europe from literature as well as precipitation range would be useful.  One to two paragraphs with appropriate citations will help. There are a handful of N. American papers that discuss geographic distribution of groundwater-fed mires [fens] as well. Some literature on how extensive these peatlands are in temperate and boreal landscapes would be helpful to convey the relevance of this work.  Some more description on differences in source hydrology to the sites studied and implications for how that may translate to strength of thermal buffer is warranted. Strictly speaking source hydrology – dominated by groundwater flows vs. hydrology dominated moreso by precipitation should have an effect on the microclimate thermal buffer. Some of the variation among sites that is unexplained by elevation may be a result of stronger vs. weaker connection to groundwater. This is an angle you really don’t address in much detail and is a potentially really important area to explore in discussion. It’s possible that wetlands in similar elevation and latitude could have markedly differing responses to climate change. i.e., stronger buffer effects at certain sites vs. others at the similar elevations – explore and describe these results further.Including a table with shift in climate trend per decade for precipitation and temperature for the study area would be informative or a citation with similar information. Some additional historical background on groundwater thermal buffer effect would improve the paper, and further support the notion that these effects have been “known” but little quantified. The authors will probably enjoy reading Clara May Frederick’s, 1974 A Natural History Study of the Vascular Flora of Cedar Bog, Champaign County, Ohio) Ohio Journal of Science who performed early temperature data collection work in such settings as far back as the 1960’s.A Ph.D. dissertation chapter by Raney (2014) Identifying Potential Refugia from Climate Change in Wetlands: High Resolution Soil Climate Maps For Priority Conservation Areas: Hydrologic Processes Mediate Significant Microclimate Variation uses very similar field data collection  methods to test the groundwater thermal buffer effect in wetlands as the authors, but focused on spatial coverage vs temporal length. That study includes 1-year worth of data. While the Raney study has not undergone 
formal peer-review and has some limitations, the soil temperature figures and data summaries in that study may be of particular interest to the authors as additional evidence of thermal buffer due to groundwater. Line 39 – increasingly there are downscaling efforts based on atmospheric microclimate networks distributed in complex terrain that focus on topographic effects on air temperature. So to a degree microclimate factors are increasingly considered, but not hydrologic factors under edaphic/soil control. It may be a distinction worth pointing out.Lines 47-59 The term mire may vary somewhat in terms of meaning by region. Please clearly define the primary source of hydrology along with your definition of mire. This will help readers in other regions. If the authors have soil-pH or pore-water pH available for study sites that would be useful but not mandatory to include in Table 1.Line 56 – insert “across their range” after homogeneous floraLine 76 – and carbon.Line 78 – replace the word “surface” with “area”Figures: Your first figure should be a site map showing regional context, and if possible topographic gradient of your sites. Figure 1 Make all labels bold, and text larger. To reduce space, put Site name labels in the light gray plot space and reduce the resulting blank space between top and bottom panels. This figure will likely take up ½ the size when published, so labels need to be larger/darker to maintain legibility.  Also you might add to your figure caption that the amplitude of diurnal temperature range in dry sites was reduced with increasing elevation, where the thermal buffer effect was also less apparent. Any ideas why La Recoleta at higher elevation shows a really strong disparity between wet and dry locations moreso than your other high elevation sites? Do you think aspect or some other factor had an effect? May be worth exploring in your results section a little further. Line 80 – Groundwater flows are somewhat decoupled from high frequency variation in precipitation, but are still linked to lower-frequency long-term changes in precipitation, groundwater recharge and flow to some degree. Focusing some attention on this phenomenon and possibly your discussion may be useful. Just because there is an observed thermal buffer effect now does not mean it will persist indefinitely, especially if climate becomes more arid – distribution of groundwater flows supporting mires may diminish in frequency, quantity, resulting in periodic drying and even potential loss of this buffer effect. For instance, regions that are wetting and expected to become wetter based on IPCC or other projections may have more stability in the thermal buffer effect vs. drying regions. It would be beneficial to focus some of your discussion on this topic.Line 86 change representativeness to representation for simplicity.
The materials and methods section needs to be much more robust in terms of explaining the statistical platform used to perform the analyses, the packages used, citations for those platforms and packages. Why the analyses are appropriate. There’s really not much of a description of how linear modeling was performed, assumptions etc. A linear mixed-model approach would make a lot of sense instead of a simple linear model, grouping wet and dry locations by a “Site” factor to control for latent processes. Need details on sample sizes in your methods.Lines 99-101 Also include quantitative descriptors of climate in the study area (mm precip/year min and max temp/year). Line 110 – provide sampling resolution of loggers. A description of why was 5cm depth chosen is needed. Please include a map of the study sites with an inset to show regional/continental position. Lines 123 – 133 – some discussion on why these are biologically or environmentally meaningful statistics is needed. Are species distributions controlled by winter minimums? Carbon dynamics by warm temperatures? Statistical analysis:Some simple statistics on precipitation and temperature trends over the last ~hundred years in this study region would be helpful to broaden discussion regarding future uncertainty of microclimates. That would also help to put your study into perspective / macroclimatic context. Are the elevational trends we’re seeing shown for a warming region? A simple trend analysis would be useful. Mm precipitation shift/decade and °C/decade would be useful metrics to include as well as the significance of those shifts. It is a stronger argument if you can show that there’s a thermal buffer effect detectable in a warming or drying climate. Dendrochronology literature frequently includes analytical methods for such trend analyses. Results: More interpretation of the results is needed. Line 155 If you’re only looking at wet vs dry sites, wouldn’t you expect wet sites to have less variability even if the source was not groundwater? Line 167 – Delete the word prove. The evidence required to prove something is very high. Suggest changing to: The results presented provide strong evidence of... Line 173 – Raney 2014 has similar findings from year-round data for a single year. Warmer in winter, cooler in summer. You should highlight that your findings were extended to a full year and the pattern is 
reproducible among years out to five years. The fact that your data possess a strong-elevational relationship, i.e., stronger under warming environment is worth more attention. Line 194 – Replace “highness” with “height”Line 200 – you elude to some topographic microclimate effects on your sites. That is fine as they influence both wet and dry locations equally, but may be worth including aspect measurements for your sites in Table 1. Neat datasets, I look forward to reading the final manuscript.