### 2020-08-29 README.Rmd
Revert "readme picture"
This reverts commit 64459c81772228a298a7502651cf5307ff5ceb57.

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![Alkaline fen in the Cantabrian Mountains](data/Sierra Cebolleda.JPG)

# Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils

This repository stores all the information related to the publication *Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in the folders of this repository:

* `data` Original raw files including soil temperature recording.
* `doc` Files to create the manuscript using `Rmarkdown`.
* `results` Output of the `R` scripts, including clean data and figures.
* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.

## Citation

Please cite the dataset as `CITATION`