### 2020-10-11 doc/Supplementary.Rmd

---
title: "Supporting information for Mire microclimate: groundwater buffers temperature in waterlogged versus dry soils"
output: 
  bookdown::word_document2:
    reference_docx: ijctemplate.docx
bibliography: refs.bib
biblio-style: international-journal-of-climatology.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(flextable); library(lubridate)
load("../results/logs.RData")

# 2015-2016 only for analysis

temperatures %>%
  filter(Time >= as.Date("2015-01-01")) %>%
  filter(Time <= as.Date("2017-12-31")) %>%
  filter(! Logger %in% c("A511BE", "A511C1")) -> temperatures

# Calculations

## Bioclimatic variables as defined by US Geological Survey, WorlClim

### Bio 2 Annual Mean Diurnal Range

temperatures %>%
  group_by(Logger, Day = floor_date(Time, "day")) %>%
  summarise(Tmax = max(Temperature), 
            Tmin = min(Temperature),
            Tdif = Tmax - Tmin) %>%
  group_by(Logger) %>%
  summarise(`Diurnal range` = mean(Tdif)) -> bio2

### Bio 5 Max Temperature of Warmest Month

temperatures %>%
  group_by(Logger, Day = floor_date(Time, "day")) %>%
  summarise(Tmax = max(Temperature)) %>%
  mutate(Month = month(Day)) %>%
  group_by(Logger, Month) %>%
  summarise(`Summer max` = mean(Tmax)) %>%
  group_by(Logger) %>%
  filter(`Summer max` == max(`Summer max`)) %>%
  select(-Month) -> bio5

### Bio 6 Min Temperature of the Coldest Month

temperatures %>%
  group_by(Logger, Day = floor_date(Time, "day")) %>%
  summarise(Tmin = min(Temperature)) %>%
  mutate(Month = month(Day)) %>%
  group_by(Logger, Month) %>%
  summarise(`Winter min` = mean(Tmin)) %>%
  group_by(Logger) %>%
  filter(`Winter min` == min(`Winter min`)) %>%
  select(-Month) -> bio6

### Merge and calculate Bio 7 Annual Temperature Range

bio2 %>%
  merge(bio5) %>%
  merge(bio6) %>%
  mutate(`Annual range` = 
           `Summer max` -
           `Winter min`) -> bioclimatics

# Tables

## Paired T-tests

t.testEFP <- function(df) 
{
  t <- t.test(df$Waterlogged, df$Dry, alternative = "less", paired = TRUE)
  data.frame(t$statistic,
             t$p.value,
             t$estimate,
             t$conf.int[1],
             t$conf.int[2]
  )
}

t.testEFPgreater <- function(df) 
{
  t <- t.test(df$Waterlogged, df$Dry, alternative = "greater", paired = TRUE)
  data.frame(t$statistic,
             t$p.value,
             t$estimate,
             t$conf.int[1],
             t$conf.int[2]
  ) # In the case of the minimum temperature is the opposite hypothesis
}

rbind(
  merge(header[, 1:8], bioclimatics, by = "Logger") %>% 
    gather(Trait, Value, `Diurnal range`:`Annual range`) %>%
    dplyr::select(Site, Groundwater, Value, Trait, Value) %>%
    spread(Groundwater, Value) %>%
    filter(! Trait %in% "Winter min") %>%
    group_by(Trait) %>%
    do(t.testEFP(.)),
  merge(header[, 1:8], bioclimatics, by = "Logger") %>% 
    gather(Trait, Value, `Diurnal range`:`Annual range`) %>%
    dplyr::select(Site, Groundwater, Value, Trait, Value) %>%
    spread(Groundwater, Value) %>%
    filter(Trait %in% "Winter min") %>%
    group_by(Trait) %>%
    do(t.testEFPgreater(.))) -> ttests

## Table 3 - Chelsa

bioclimatics %>%
  gather(Variable, Value, `Diurnal range`:`Annual range`) %>%
  mutate(Source = "Loggers") -> bioLoggers

header %>%
  select(Logger, `Diurnal range`:`Annual range`) %>%
  gather(Variable, Value, `Diurnal range`:`Annual range`) %>%
  mutate(Source = "CHELSA") -> bioCHELSA

header %>%
  select(Site, Logger, Groundwater, Mire) %>%
  merge(rbind(bioLoggers, bioCHELSA), by = "Logger") %>%
  spread(Source, Value) %>%
  na.omit -> bio

RMSE <- function(df) {
  require(Metrics)
  data.frame(RMSE = rmse(df$Loggers, df$CHELSA))}

CORR <- function(df) {
  c <- cor.test(df$Loggers, df$CHELSA)
  data.frame(t = c$statistic,
             df = c$parameter,
             p = c$p.value,
             r = c$estimate)}

bio %>%
  group_by(Variable, Groundwater) %>%
  do(CORR(.)) -> corrtable # Pearson's correlation

bio %>%
  group_by(Variable, Groundwater) %>%
  do(RMSE(.)) -> rmsetable # Root-mean-square error (RMSE)

merge(corrtable, rmsetable) %>%
  mutate(RMSE = round(RMSE, 2),
         r = round(r, 2),
         t = round(t, 3),
         p = round(p, 3)) %>%
  rename(`RMSE (ºC)` = RMSE,
         `Pearson's r` = r) -> chelsa
```

```{r table1, echo = FALSE}
regulartable(ttests) %>% width(width = 1) %>% set_caption("Table S1: Results of the t-tests comparing the paired waterlogged-dry soil bioclimatic variables. Analysis performed with full-year, contemporaneous data series (2015-2017).")
```

```{r table3, echo = FALSE}
regulartable(chelsa) %>% width(width = 1) %>% set_caption("Table S2: Results of the linear models predicting the soil bioclimatic variables from the CHELSA air temperatures, per bioclimatic variable and groundwater situation. RMSE is the root-mean-square error, a measure of the accuracy of the predictions (lower values indicate higher accuracy). Analysis performed with full-year, contemporaneous data series (2015-2017).")
```